### 2014-03-31 paper_enahpc14/Introduction.tex
updates paper; adds two empty files for the redaction of COSMO-ART model description and setup.

\noindent 
Aerosols are  suspensions of liquid,  solid or mixed particles  in the
air  (including sea  salt,  mineral dust,  sulphate, nitrate,  organic
carbon and  black carbon),  with highly variable  chemical composition
and  size  distribution \citep{Putaud-2010}.   Although  they are  not
considered a heat-trapping greenhouse gas and have shorter atmospheric
lifetimes, they significantly affect  the Earth's radiation budget and
are acknowledged as one of  the most significant and uncertain aspects
in  anthropogenic forcing  over the  last 150  years \citep{Koch-2009,
IPCC-2013}.  Enhanced  aerosols concentrations can  impact the climate
system by  reflecting (e.g.  pure sulfates and  nitrates) or absorbing
(e.g.  black  carbon) solar radiation  and thereby exert a  cooling or
warming  effect on  the Earth-atmosphere  system, causing  a so-called
direct    radiative   forcing    \citep{Charlson-1991,   Haywood-2000,
Ramanathan-2001,  Liao-2005, Bangert-2012,  Lundgren-2013}.  Depending
on their  size and  chemical composition, they  can also act  as cloud
condensation  and   ice  nuclei   and  profoundly  impact   the  cloud
microphysical   processes  and   optical   properties  (cloud   albedo
effect,   \citealp{Twomey-1977}),  leading   to  changes   in  droplet
concentrations                \citep{Albrecht-1989}                and
precipitation   \citep{Rosenfeld-2000,   Khain-2008,  Pruppacher-2010,
Seifert-2012, Tao-2012, Lee-2012},  and therefore producing a negative
indirect   radiative    forcing   \citep{Haywood-2000,   Lohmann-2005,
VandenHeever-2011,   Rosenfeld-2013}.    However,   the   effects   of
anthropogenic aerosols on clouds and the hydrological cycle as well as
the  cloud   lifetime  effect  are  especially  hard   to  assess  and
quantify   \citep{IPCC-2013},   and   remain   one  of   the   largest
uncertainties in climate modeling and in climate change prediction due
to the lack of  understanding of cloud feedbacks \citep{Sherwood-2013,
Rosenfeld-2013, Lee-2013}. \\

\noindent 
Hence to  enhance our understanding of  aerosol-cloud interactions and
reduce  uncertainties  of aerosol  effects  in  climate, the  research
community is making a  concerted international effort to represent the
underlying chemical  processes through models.  These  models, such as
ART (Aerosols  and Reactive Trace  gases) extension, developed  at the
Karlsruhe Institute  of Technology (KIT),  offer a key  opportunity to
reduce   the  climate  uncertainty,   particularly  on   the  regional
scale \citep{Knote-2011, Bangert-2011,  Knote-2013}.  These models can
be coupled with climate models,  such as the regional weather forecast
model COSMO (Consortium  for Small-scale Modelling), jointly developed
by  a consortium  of  European weather  centers  including the  German
weather service  DWD and MeteoSwiss,  and used in the  climate version
COSMO-CLM by a wide  research community.  The extended COSMO-ART model
provides a detailed description of air pollution chemistry and aerosol
processes, and  is mainly  designed to study  air quality  and aerosol
meteorology feedbacks on short, episodic to annual time scales.  It is
capable  of   simulating  aerosol  distributions  as   well  as  their
interactions over Europe as well as other regional domains.\\

\noindent 
COSMO-ART is  computationally much more demanding than  the COSMO core
since a  large number of additional  tracers and processes  have to be
considered.  Thus this model is currently severely-limited in terms of
applicability and expensive in terms of energy consumption.  COSMO has
recently  been  ported  to  GPUs  within the  framework  of  the  High
Performance  and  High  Productivity  Computing (HP2C)  Initiative  to
optimize it  for computational and energy  efficiency.  Although these
developments will  facilitate the  application of COSMO  for numerical
weather prediction and climate  simulations, they do little to address
the  coupling with  the  ART model  extension,  for which  significant
investments  are still required  to take  it to  a similar  level. The
efficiency of ART is being  addressed in the EU Exa2Green project.  In
Sec.~\ref{sec:1},  we   introduce  the  aerosol  model   and  its  its
integration  into  COSMO.    Sec.~\ref{sec:2}  presents  the  hardware
considerations.  In  Sec.~\ref{sec:3} we describe  measurement devices
and energy-aware techniques employed  to evaluate the energy footprint
of the  considered application.  Finally,  Sec.~\ref{sec:4} presents a
benchmark  analysis  of  COSMO-ART   from  the  Exa2Green  project  to
highlight the critical components and an outline of the next steps for
model optimization.  The ultimate goal  of the project is to deliver a
prototype code, which  provides an energy efficiency of  at least five
times of the  baseline value.  Such an implementation  would allow the
community to  investigate critical questions at  higher resolution and
over longer periods, at reduced cost to the environment.

>>>>

\noindent 
Aerosols are  suspensions of liquid,  solid or mixed particles  in the
air  (including sea  salt,  mineral dust,  sulphate, nitrate,  organic
carbon and  black carbon),  with highly variable  chemical composition
and  size  distribution \citep{Putaud-2010}.   Although  they are  not
considered a heat-trapping greenhouse gas and have shorter atmospheric
lifetimes, they significantly  affect the atmospheric radiative fluxes
and  are acknowledged  as one  of the  most significant  and uncertain
aspects    in    anthropogenic    forcing    over   the    last    150
years \citep{Koch-2009,  IPCC-2013}.  Enhanced aerosols concentrations
can impact the  climate system by reflecting (e.g.   pure sulfates and
nitrates)  or  absorbing  (e.g.   black carbon)  solar  radiation  and
thereby  exert a  cooling or  warming effect  on  the Earth-atmosphere
system,      causing       a      so-called      direct      radiative
forcing     \citep{Charlson-1991,    Haywood-2000,    Ramanathan-2001,
Liao-2005, Bangert-2012, Lundgren-2013}.   Depending on their size and
chemical composition, they can also  act as cloud condensation and ice
nuclei, and  thereby influence  the cloud microphysical  processes and
optical   properties  (cloud  albedo   effect,  \citealp{Twomey-1977})
through the contribution to  cloud formation.  This results in changes
in      droplet      concentrations     \citep{Albrecht-1989}      and
precipitation     \citep{Rosenfeld-2000,    Khain-2008,    Khain-2009,
Pruppacher-2010,   Seifert-2012,  Tao-2012,  Lee-2012},   producing  a
so-called  negative  indirect  radiative forcing  \citep{Haywood-2000,
Lohmann-2005,  VandenHeever-2011, Rosenfeld-2013}.  Wet  scavenging of
aerosol  particles  represents  a  major  removal  mechanism  for  air
pollutants from  the atmosphere, however the  effects of anthropogenic
aerosols on  clouds and  the hydrological cycle  as well as  the cloud
lifetime    effect    are    especially    hard    to    assess    and
quantify   \citep{IPCC-2013},   and   remain   one  of   the   largest
uncertainties in climate modeling and in climate change prediction due
to  the lack  of  knowledge in  cloud feedbacks  \citep{Sherwood-2013,
Rosenfeld-2013, Lee-2013}. \\

\noindent 
Hence to  improve our understanding of  aerosol-cloud interactions and
reduce  uncertainties  of aerosol  effects  in  climate, the  research
community is making a  concerted international effort to represent the
underlying chemical  processes through models.  These  models, such as
ART (Aerosols  and Reactive Trace  gases) extension, developed  at the
Karlsruhe Institute  of Technology (KIT),  offer a key  opportunity to
reduce   the  climate  uncertainty,   particularly  on   the  regional
scale \citep{Knote-2011, Bangert-2011,  Knote-2013}.  These models can
be coupled with climate models,  such as the regional weather forecast
model COSMO (Consortium  for Small-scale Modelling), jointly developed
by  a consortium  of  European weather  centers  including the  German
weather service  DWD and MeteoSwiss,  and used in the  climate version
COSMO-CLM by a wide  research community.  The extended COSMO-ART model
provides a detailed description of air pollution chemistry and aerosol
processes, and  is mainly  designed to study  air quality  and aerosol
meteorology feedbacks on short, episodic to annual time scales.  It is
capable  of   simulating  aerosol  distributions  as   well  as  their
interactions over Europe as well as other regional domains.\\

\noindent 
COSMO-ART is  computationally much more demanding than  the COSMO core
since a  large number of additional  tracers and processes  have to be
considered.  Thus this model is currently severely-limited in terms of
applicability and expensive in terms of energy consumption.  COSMO has
recently  been  ported  to  GPUs  within the  framework  of  the  High
Performance  and  High  Productivity  Computing (HP2C)  Initiative  to
optimize it  for computational and energy  efficiency.  Although these
developments will  facilitate the  application of COSMO  for numerical
weather prediction and climate  simulations, they do little to address
the  coupling with  the  ART model  extension,  for which  significant
investments  are still required  to take  it to  a similar  level. The
efficiency of ART is being  addressed in the EU Exa2Green project. The
ultimate goal  of the  project is to  deliver a prototype  code, which
provides an energy  efficiency of at least five  times of the baseline
value.   Such   an  implementation   would  allow  the   community  to
investigate critical  questions at  higher resolution and  over longer
periods, at reduced cost  to the environment.  In Sec.~\ref{sec:1}, we
introduce the online coupled  mesoscale atmosphere and chemistry model
COSMO-ART  \citep{Vogel-2009}  that  accounts  for  feedbacks  between
chemistry,  aerosols,  radiation,  and clouds.   Sec.~\ref{sec:2}  and
Sec.~\ref{sec:3}  successively describe  the hardware  and measurement
devices used in evaluation as well as energy-aware techniques employed
to  determine  the energy  footprint  of  the considered  application.
Sec.~\ref{sec:4}  shows benchmark results  from the  Exa2Green project
and highlight the critical  components of the application.  Finally, a
summary   and  directions   for  model   optimization  are   given  in
Sec.~\ref{concl}.
